Mining the human microbiome for clinical insight

### 2018-12-15 abstract.tex
Prep to add 24hr and perc norm chapters

The human microbiome is essential for health and has been implicated in many diseases.
DNA sequencing has enabled the detailed characterization of these human-associated microbial communities, leading to a rapid expansion in studies investigating the human microbiome.
%However, extracting clinically-relevant associations from microbiome datasets remains challenging because of the high dimensional nature of the data and variability across studies.
In this thesis, I describe three projects which overcome various data analysis challenges to extract useful clinical insights from microbiome data.
In the first project, I present an analysis of lung, stomach, and oropharyngeal microbiomes.
I leverage data collected from multiple sites per patient to identify aspiration-associated changes in the relationships between aerodigestive communities, discovering new properties of the aerodigestive microbiome and suggesting new approaches for treatment.
%These changes suggest new approaches for developing diagnostics and treatments for aspiration-related respiratory complications.
%I leverage data collected from multiple sites per patient to find aspiration-associated changes in the relationships between aerodigestive communities which suggests new targets for treatment and diagnosis.
In the second project, I perform a meta-analysis of case-control gut microbiome datasets with standard data processing and analysis methods.
I find consistent patterns characterizing disease-associated microbiome changes and a set of shared associations which could inform clinical treatment and therapeutic development approaches for different microbiome-mediate diseases.
In the third project, I describe a framework for rational donor selection in fecal microbiota transplant clinical trials.
In this framework, knowledge derived from clinical and basic science research is used to inform which donor is selected for fecal transplants, increasing the likelihood of successful trials.
Together, these projects demonstrate a variety of approaches to mine the human microbiome for clinically-relevant insights and suggests multiple avenues forward for translating findings from microbiome data analyses into clinical impact.

>>>>

The human microbiome is essential for health and has been implicated in many diseases.
DNA sequencing has enabled the detailed characterization of these human-associated microbial communities, leading to a rapid expansion in studies investigating the human microbiome.
In this thesis, I describe multiple projects which overcome various data analysis challenges to extract useful clinical insights from microbiome data.
In the first project, I present an analysis of lung, stomach, and oropharyngeal microbiomes.
I leverage data collected from multiple sites per patient to identify aspiration-associated changes in the relationships between aerodigestive communities, discovering new properties of the aerodigestive microbiome and suggesting new approaches for treatment.
In the second project, I perform a meta-analysis of case-control gut microbiome datasets with standard data processing and analysis methods.
I find consistent patterns characterizing disease-associated microbiome changes and a set of shared associations which could inform clinical treatment and therapeutic development approaches for different microbiome-mediate diseases.
Enabled by this work, I also contribute to the development of a non-parametric method to correct for batch effects in case-control microbiome studies.
In the fourth project, I describe a framework for rational donor selection in fecal microbiota transplant clinical trials.
In this framework, knowledge derived from clinical and basic science research is used to inform which donor is selected for fecal transplants, increasing the likelihood of successful trials.
Finally, I present preliminary results analyzing the microbiome and metabolome of residential sewage towards community-level public health surveillance.  
Together, these projects demonstrate a variety of approaches to mine the human microbiome for clinically-relevant insights and suggests multiple avenues forward for translating findings from microbiome data analyses into clinical and public health impact.

### 2018-12-15 chap_24hr.tex
Prep to add 24hr and perc norm chapters

\graphicspath{{donor-selection/figures/}}

\chapter{Increasing fecal microbiota transplant clinical trial successes by applying rational donor selection}

\noindent
Claire Duvallet, Caroline Zellmer, Pratik Panchal, Shrish Budree, Majdi Osman, and Eric J Alm.

\bigskip
\noindent
The contents of this chapter constitute a manuscript in preparation for submission.

\bigskip
\input{donor-selection/ms.tex}

>>>>

%\graphicspath{{donor-selection/figures/}}

\chapter{Residential sewage provides a view into community-level human microbiome and metabolome}

\noindent
Mariana G Matus, Claire Duvallet, Melissa Soule, Noriko Endo, Newsha Ghaeli,  Ilana Brito, Carlo Ratti, Elizabeth B Kujawinski, Eric J Alm

\bigskip
\bigskip
\noindent
The contents of this chapter constitute a manuscript in preparation for submission.

\clearpage
%\input{donor-selection/ms.tex}

### 2018-12-15 chap_aspiration.tex
Prep to add 24hr and perc norm chapters

The contents of this chapter have been submitted for publication as:

Claire Duvallet, Kara Larson, Scott Snapper, Sonia Iosim, Ann Lee, Katherine Freer, Kara May, Eric J. Alm, and Rachel Rosen. "Aerodigestive sampling reveals altered microbial exchange between lung, oropharyngeal, and gastric microbiomes in children with impaired swallow function." In review, November 2018.

The tables and figures are at the end of the Chapter. The Supplementary Information is in Appendix A.

>>>>

Claire Duvallet, Kara Larson, Scott Snapper, Sonia Iosim, Ann Lee, Katherine Freer, Kara May, Eric J Alm, and Rachel Rosen.

\bigskip
\bigskip
\noindent
The contents of this chapter have been submitted for publication as: ``Aerodigestive sampling reveals altered microbial exchange between lung, oropharyngeal, and gastric microbiomes in children with impaired swallow function,'' in review as of December 2018.

\bigskip
\bigskip
\noindent
The tables and figures are at the end of the Chapter. The Supplementary Information is in Appendix A.

### 2018-12-15 chap_conclusion.tex
Prep to add 24hr and perc norm chapters

A turning point in my thesis came when I read the 1976 paper coining the term "meta-analysis" \cite{glass-1976}.
In it, Glass argues for the importance and underappreciated value of consolidating and synthesizing information into knowledge, prizing work that aims to find meaning and draw conclusions from existing studies.
This was a turning point for two reasons.
First, it was the moment I became truly proud of my work, especially the meta-analysis: I realized that it was not just some sort of microbiome "stamp collection" endeavor that anyone with basic knowledge of computational tools could do, but was instead difficult and valuable work that I uniquely contributed to.
Second, it helped me recognize a uniting theme in all of my work as exactly what Glass wrote, "find[ing] the knowledge in the information."
I realized that in each analysis in this thesis, I aimed to move beyond statistical associations all the way to clinical implications.
I hope that as we move forward in this exciting and vibrant field, we collectively become less satisfied with simply reporting new information and emphasize instead efforts to synthesize existing knowledge and generate new insights that lead more directly to clinical impact.

>>>>

A turning point in my thesis came when I read Gene Glass's 1976 paper coining the term ``meta-analysis''"'' \cite{glass-1976}.
In it, Glass argues for the underappreciated importance of consolidating and synthesizing information into knowledge, prizing work that aims to find meaning and draw conclusions from disparate existing studies.
This was a turning point in my thesis for two reasons.
First, it was the moment I became truly proud of my work, especially the meta-analysis in Chapter 3: I realized that it was not just some sort of microbiome ``stamp collection'' endeavor that anyone with basic knowledge of computational tools could do, but was instead difficult and valuable work that I had uniquely contributed to.
Second, reading Glass's words helped me recognize a uniting theme in all of my work: ``find[ing] the knowledge in the information.''
I realized that in each project presented in this thesis, I had not just mined the data to find statistical associations, but rather to interpret associations and chase them all the way to their potential clinical implications.
I hope that as we move forward in this exciting and vibrant field, microbiome researchers collectively become less satisfied with simply reporting new information, instead emphasizing and valuing efforts that synthesize existing knowledge and generate new insights that lead more directly to clinical impact.

### 2018-12-15 chap_meta-analysis.tex
Prep to add 24hr and perc norm chapters

The contents of this chapter were published as:

Claire Duvallet, Sean M. Gibbons, Thomas Gurry, Rafael A. Irizarry, and Eric J. Alm. "Meta-analysis of gut microbiome studies identifies disease-specific and shared responses." \textit{Nature Communications}. (2017)

The figures are at the end of the chapter. The Supplementary Information is in Appendix B.

>>>>

Claire Duvallet, Sean M Gibbons, Thomas Gurry, Rafael A Irizarry, and Eric J Alm.

\bigskip
\bigskip
\noindent
The contents of this chapter were published as: ``Meta-analysis of gut microbiome studies identifies disease-specific and shared responses'' in \textit{Nature Communications} (2017) volume 8, article number 1784.

\bigskip
\bigskip
\noindent
The figures are at the end of the chapter. The Supplementary Information is in Appendix B.

### 2018-12-15 chap_perc-norm.tex
Prep to add 24hr and perc norm chapters

\chapter{Residential sewage provides a view into community-level human microbiome and metabolome}

\noindent
Mariana G Matus, Claire Duvallet, Melissa Soule, Noriko Endo, Newsha Ghaeli,  Ilana Brito, Carlo Ratti, Elizabeth B Kujawinski, Eric J Alm

\bigskip
\bigskip
\noindent
The contents of this chapter constitute a manuscript in preparation for submission.

>>>>

\chapter{Correcting for batch effects in case-control microbiome studies}

\noindent
Sean M Gibbons, Claire Duvallet, and Eric J Alm

\bigskip
\bigskip
\noindent
The contents of this chapter were published as: ``Correcting for batch effects in case-control microbiome studies'' in \textit{PLoS Computational Biology} (2018) volume 14(4), page e1006102.

### 2018-12-15 cover.tex
Prep to add 24hr and perc norm chapters

Mentors:
Rafa - will always be a friend of the lab
Jim
Deb
Ilana - for being the mentor I didn't know I needed

>>>>

Mentors:
Rafa - will always be a friend of the lab
Jim
Deb
Ilana - for being the mentor I didn't know I needed
Liz - teaching me everything I know about metabolomics and being so giving with time
